---
title: "spatialSpells: Usage"
output: 
  rmarkdown::html_document:
    toc: true
    toc_title: "Content"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval      = TRUE,
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 8,
  fig.align = "center",
  dpi = 150,
  out.width = "100%"
)
```


## Overview

This article shows usage examples of the `infra` functions in the R package `spatialSpells`. The prefix `infra` identifies low level functions designed as infrastructure for other packages of the `spatial` family. This article 


## Setup

The code below loads the `R` packages and example data required for this tutorial.

```{r, include = FALSE}
if (!requireNamespace("pak", quietly = TRUE)) {
  install.packages("pak")
}
library(pak)
pak::pkg_install(
  c(
    "dplyr",
    "sf",
    "mapview",
    "terra",
    "viridis",
    "spatialData"
  ),
  ask = FALSE
)
```

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(sf)
library(mapview)
library(terra)
library(viridis)
library(spatialData)
library(spatialSpells)
```

## Example Data

Here we prepare some example datasets to show how `spatialSpells` works.

The code below prepares a data frame with coordinates.

```{r}
df_coordinates <- spatialData::vi |> 
  dplyr::mutate(
    longitude = sf::st_coordinates(geometry)[, "X"],
    latitude = sf::st_coordinates(geometry)[, "Y"]
  ) |> 
  sf::st_drop_geometry() |> 
  dplyr::select(
    vi_numeric,
    longitude,
    latitude
  )

df_coordinates |> 
  head() |> 
  dplyr::glimpse()
```
In a similar way, the code below prepares a dataframe with polygons coded a well-known text (WKT). This one will take a minute, it's downloading a large-ish dataset.

```{r}
df_wkt <- spatialData::plantae_extra() |> 
  dplyr::select(
    ecoregion_id, 
    richness_species, 
    geometry
    ) |> 
  dplyr::mutate(
    wkt = sf::st_as_text(geometry)
  ) |> 
  sf::st_drop_geometry()

df_wkt |> 
  head() |> 
  dplyr::glimpse()
```



### Identification of columns with coordinates

The datasets `infra_x_column_names` and `infra_y_column_names` contain common names for columns with coordinates in several languages.

```{r}
infra_x_column_names
```
```{r}
infra_y_column_names
```
These obects
